---
title: "Course assignment"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
An assignment to understand R markdown and reproducible research.
Libraries that will be used.
```{r}
library(dplyr)
library(ggplot2)
```

## Assignment

The first step is to load the data set that we will be working with.
```{r}
activity <- read.csv("data/activity.csv")
activity <- tibble::as_tibble(activity)
head(activity, n = 10)
```

Histogram of the total steps in a day.
```{r}
by_sum <- activity %>% group_by(date) %>% summarise(sum_Steps = sum(steps), na.rm = TRUE)
by_sum <- tibble::as_tibble(by_sum)



ggplot(data = by_sum, aes(x = sum_Steps)) + geom_histogram (bins = 6, fill = "red", col = "black", alpha = .5) + labs (x = "Steps", y = "Frequency", title = "Histogram of Total steps")

```
Mean and Median for the steps taken each day.
```{r}
summary(by_sum)
```
The median is presented as 10,765 as for the mean is 10,766.

Time series plot of the 5-minute interval (x-axis) and the average steps per day (y-axis).
```{r}
avg_interval <- activity %>% group_by(interval) %>% summarize(avg_steps = mean(steps, na.rm = TRUE))
qplot(interval, avg_steps, data = avg_interval, geom=c("line"), xlab = "Interval", ylab = "Average steps per day", main = "Time Series plot")
```
Finding the number of missing values in the data set.
```{r}
sum(is.na(activity))
```
Removing the missing values from the data using the mean average steps per day.
```{r}
for (i in 1:nrow(activity)) {
    
    if(is.na(activity$steps[i])) {
        
        val <- avg_interval$avg_steps[which(avg_interval$interval == activity$interval[i])]
        activity$steps[i] <- val 
   }
}

sum(is.na(activity))

```
Histogram with the missing value replaced.
```{r}
by_sum <- activity %>% group_by(date) %>% summarise(sum_Steps = sum(steps))
by_sum <- tibble::as_tibble(by_sum)



ggplot(data = by_sum, aes(x = sum_Steps)) + geom_histogram (bins = 6, fill = "red", col = "black", alpha = .5) + labs (x = "Steps", y = "Frequency", title = "Histogram of Total steps")
```
Lets see if the mean and median have changed.
```{r}
summary(by_sum)
```

Lastly, we will try to see if there is a pattern between steps in the weekday and steps in the weekend.

```{r}
activity['type_of_day'] <- weekdays(as.Date(activity$date))

activity$type_of_day[activity$type_of_day  %in% c('Saturday','Sunday') ] <- "weekend"
activity$type_of_day[activity$type_of_day != "weekend"] <- "weekday"

activity$type_of_day <- as.factor(activity$type_of_day)
interval_sum <- aggregate(steps ~ interval + type_of_day, activity, mean)

qplot(interval, steps, data = interval_sum, geom=c("line"), xlab = "Interval", ylab = "Number of steps", main = "") + facet_wrap(~ type_of_day, ncol = 1)
```